Deciphering Key Determinants of Biden Voter Support: Insights from Logistic Regression Analysis in the 2020 US Presidential Election

### 2024-03-17 paper/paper.qmd
Paper Updates

---
title: "Deciphering Key Determinants of Biden Voter Support: Insights from Logistic Regression Analysis in the 2020 US Presidential Election"
author: 
  - Fares Alkorani
  - Owais Zahid
thanks: "Code and data are available at: https://github.com/FFFiend/linear_model_investigation."
date: today
date-format: long
abstract: "On November 4, 2020, the 59th United States presidental election took place where Joe Biden became elected as 46th president of the United States. Using responses from the CES 2020 election survey (n = 8000), we study the effect that gender, race, age, and education have on whether an American voted for Joe Biden using a logistic regression model. We find that men are less likely to vote for Biden than women, respondents with higher education level are more likely to vote for Biden, and that Black respondents are most likely to vote for Biden from all other racial groups. We also find a weak negative correlation between a respondents age and whether they vote for Biden. As the upcoming United States presidental election is anticipated to be between Donald J. Trump and Joe Biden, these results have implications for how both candidates will use voter demographics to guide their campaigns."
format: pdf
number-sections: true
bibliography: references.bib
---
# Introduction

@@

As a result, in this paper the estimand we hope to explore is the likelihood that an American citizen will vote for Joe Biden in the 2020 presidental election given their income, gender, race, and age. We will do this by using the 2020 Cooperative Election Study (CES) [CITATION] which is a national online survey that was administered to a representative sample of American citizens pre and post-election, gathering demographic information from 61,000 respondents. Using 8000 responses from this survey, we construct a logistic regression model using income, gender, race, and age as predictor variables to see which demographics aligned with support for Joe Biden. We find that (**I NEED YOUR MOST RELEVANT FINDINGS HERE**). This information is important as it can allow us to understand the campaign strategies in the 2024 United States presidential election toward citizens with specific demographics, especially as Donald Trump and Joe Biden will likely be representatives of their respective parties for the upcoming presidential election.

The remainder of this paper is structured as follows: Section 2 discusses the data from the 2020 CES, Section 3 discusses the logistic regression model that was constructed from the 2020 CES data, Section 4 presents the results, and finally Section 5 discusses our findings and some weaknesses. (**MOSTLY FROM https://tellingstorieswithdata.com/04-writing_research.html#introduction-1, DUNNO IF IT NEEDS REPHRASING OR GIVEN ADDITIONAL DETAILS LATER**).

@@

# Model
We resort to using a Logistic Regression model since we are interested in predicting the candidate a survey respondent voted for, given the former's age, gender, education and race. A logistic regression model is suitable as our response variables (1 for Biden and 0 for Trump) are clearly binary.

@@

# Results

@@

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

## Diagnostics

\newpage

>>>>

---
title: "Deciphering Key Determinants of Biden Voter Support: Insights from Logistic Regression Analysis in the 2020 US Presidential Election"
author: 
  - Fares Alkorani
  - Owais Zahid
thanks: "Code and data are available at: https://github.com/FFFiend/us_2020_election_analysis."
date: today
date-format: long
abstract: "On November 4, 2020, the 59th United States presidental election took place where Joe Biden became elected as 46th president of the United States. Using responses from the CES 2020 election survey (n = 8000), we study the effect that gender, race, age, and education have on whether an American voted for Joe Biden using a logistic regression model. We find that men are less likely to vote for Biden than women, respondents with higher education level are more likely to vote for Biden, and that Black respondents are most likely to vote for Biden from all other racial groups. We also find a weak negative correlation between a respondents age and whether they vote for Biden. As the upcoming United States presidental election is anticipated to be between Donald J. Trump and Joe Biden, these results have implications for how both candidates will use voter demographics to guide their campaigns."
format: pdf
number-sections: true
bibliography: references.bib
---
# Introduction

@@

As a result, in this paper the estimand we hope to explore is the likelihood that an American citizen will vote for Joe Biden in the 2020 presidental election given their income, gender, race, and age. We will do this by using the 2020 Cooperative Election Study (CES) [CITATION] which is a national online survey that was administered to a representative sample of American citizens pre and post-election, gathering demographic information from 61,000 respondents. Using 8000 responses from this survey, we construct a logistic regression model using income, gender, race, and age as predictor variables to see which demographics aligned with support for Joe Biden. We find that men are less likely to vote for Biden than women, respondents with higher education level are more likely to vote for Biden, and that Black respondents are most likely to vote for Biden from all other racial groups. We also find a weak negative correlation between a respondents age and whether they vote for Biden. This information is important as it can allow us to understand the campaign strategies in the 2024 United States presidential election toward citizens with specific demographics, especially as Donald Trump and Joe Biden will likely be representatives of their respective parties for the upcoming presidential election.

The remainder of this paper is structured as follows: @sec-data discusses the data from the 2020 CES, @sec-model discusses the logistic regression model that was constructed from the 2020 CES data, and @sec-results presents the results.

@@

# Model {#sec-model}
We resort to using a Logistic Regression model since we are interested in predicting the candidate a survey respondent voted for, given the former's age, gender, education and race. A logistic regression model is suitable as our response variables (1 for Biden and 0 for Trump) are clearly binary.

@@

# Results {#sec-results}

@@
